---
title: "Population of US Cities Through History"
author: "Igor Cizelj, PhD"
date: "November 7, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Population of US Cities Through History

```{r fig.width=8, fig.height=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("./data/image.PNG")
grid.raster(img)
```

## Data Set and Shiny App

- The Shiny App is deployed at [shinyapps.io](https://igorcizelj.shinyapps.io/HistUSCitiesPopulation/).

- The data set used by the App is a compilation of US historical city populations decennially between 1850 and 2010, for a selection of states. The original data set can be found [here](https://github.com/cestastanford/historical-us-city-populations), which contains all US states from 1790 to 2010.

- Use the slider to choose the Year (from 1850 to 2010, in increments of 10).

- Use the dropdown menu to choose the State (CA, TX, NY, FL, IL, PA, OH, GA, MI, or NC). 

- Given a Year and a State, the Shiny App displays the cities on the map, and color codes them based on the population.

## Original Data

```{r fig.width=8, fig.height=4,echo=TRUE}
library(reshape)
cityData <- read.csv("./data/fullData.csv",header = FALSE,
                     skip=1,stringsAsFactors = FALSE)
cityData <- cityData[,c(2:27,31,32)]
names(cityData) <- c("State", "City", "CityState",
                     as.character(seq(1790,2010,10)),"Lat","Long")    
cityData <- melt(cityData, id = c("State","City",
                                  "CityState","Lat","Long"))
names(cityData)[c(6,7)] <- c("Year","Population")
cityData$Year <- as.numeric(as.character(cityData$Year))
cityData <- na.omit(cityData)
head(cityData,n=2)
```

## Data Used by App

```{r fig.width=8, fig.height=4,echo=TRUE}
# Remove Cities with LatLong=0,0
cityData <- cityData[which(cityData$Lat!=0 | cityData$Long!=0),]
# Extract Only Data with Particular States
cityData <- cityData[which(cityData$State %in% 
                               c('CA','TX','NY','FL','IL',
                                 'PA','OH','GA','MI','NC')),]
# Extract Only Data with Year>=1850
cityData <- cityData[which(cityData$Year>=1850),]
# Extract Only Data with Population > 100
cityData <- cityData[which(cityData$Population>100),]
head(cityData,n=2)

```